---
title: "Asian American Quality of Life Project"
output:
  github_document:
---

# Links

- [Dataset Source](https://data.austintexas.gov/dataset/Final-Report-of-the-Asian-American-Quality-of-Life/hc5t-p62z)

# Project Overview
In this project, we study the impact of cultural differences on economic well-being in the Asian American population within the Austin area. Our objective is to answer the question: do cultural factors affect economic outcomes such as income bracket, employment rate, or quality of life among Asian Americans? We aim to determine how these variables influence each other and to uncover the underlying mechanisms that contribute to economic disparities within the Asian American population. By examining factors, such as ethnicity, education, religion, and belonging, we aspire to discern their interconnectedness with the economic success indicators mentioned. 

The dataset we are examining contains survey data on the fastest-growing minority group in the United States, Asian Americans. Within this dataset, the broad term includes individuals with origins in the Far East, Southeast Asia, or the Indian subcontinent. The Final Report of the Asian American Quality of Life (AAQoL) comes from a survey completed by the city of Austin, TX from August 19 to December 12, 2015 (Jang 2016, 12). It includes 2,609 observations each corresponding to an individual survey. There are 231 columns in the dataset, each representing the various questions respondents were asked to fill out when taking the survey.


# Hypothesis 1: Income Bracket v.s. Education vs. Ethnicity

We predict a variation in educational attainment amongst the different types of Asian ethnicities. We also expect to see a positive general relationship between education and income.

```{r, echo = FALSE, message=FALSE, warning=FALSE, error=FALSE}
# Import the dataset into R environment.
library(ggplot2)
library(dplyr)
library(tidyverse)
library(readr)
library(ggthemes)
Final_Report_of_the_Asian_American_Quality_of_Life_AAQoL_ <- read_csv("Final_Report_of_the_Asian_American_Quality_of_Life__AAQoL_.csv")

# Load the dataset into a data frame named 'AAQoL'
AAQoL <- data.frame(Final_Report_of_the_Asian_American_Quality_of_Life_AAQoL_)

#Removing NAs
test3 <- AAQoL %>%
  filter(!is.na(Income) & !is.na(Education.Completed) & !is.na(Ethnicity))


#Graphing education on y-axis, income bracket on x-axis, colored by ethnicity.
new_colors <- c("orange", "red", "blue", "lightblue2", "maroon")

test3 %>% filter(Ethnicity != "Other", Student != "Student", Retired != "Retired", Unemployed != "Unemployed") %>%
  ggplot(aes(x = Income, y = Education.Completed, fill = Ethnicity)) +
  geom_bar(stat = "summary", fun = mean, position = "dodge", na.rm = TRUE) +
  geom_errorbar(stat="summary", fun.data=mean_se, width=.5, position = position_dodge(width = 0.9))+
  coord_flip() +
  coord_cartesian(ylim = c(10, 18)) +
  theme_economist_white() +
  theme(axis.text.x = element_text(angle = 50, vjust = 0.5, hjust = .4, size = 9)) +
  scale_fill_manual(values = new_colors) +
labs(x = "Household Income Bracket", y = "Mean Education Completed")
```


# Hypothesis 2: Income Bracket v.s. English Proficiency 
We predict that the level of English proficiency in the Asian American population will be positively related with their income bracket.

```{R, echo = FALSE, message=FALSE, warning=FALSE, error=FALSE}

library(tidyverse)
library(extrafont)
loadfonts()


df_unfiltered <- read_csv("Final_Report_of_the_Asian_American_Quality_of_Life__AAQoL_.csv")
# Data Wrangling, 2609 observations to 1620 observations
df <- df_unfiltered %>%
  filter(`US Born` == 'No') %>% # Filter for only immigrants, 256 rows removed.
  filter(('Full Time Employment' == 'Employed full time')|('Part Time Employment' == 'Employed part time')|(Retired=='Retired')) %>%
  filter(Income != '') %>% # Filter for only response with income information, 167 rows removed
  filter('English Speaking' != '') %>% # Filter for only responses with proficiency information, 6 rows removed
  select(1,Income,'Proficiency'='English Speaking',21:22,Retired,Age) %>% # Select only relevant variables
  na.omit() # Omit all cells with N/A, 263 rows removed.
  # TO-DO
  # Ask TA About how to join the employment columns

# Redefining Proficiency as a ordinal variable
df$Proficiency <- factor(df$Proficiency, ordered = TRUE,
                         levels = c("Not at all", "Not well", "Well", "Very well"))

# Primary Visualization
df %>% ggplot(aes(x = Income, fill = Proficiency)) +
              geom_bar(position = 'fill') +
              scale_fill_brewer(palette = 15) +
              theme(axis.text.x  = element_text(angle = 15, hjust = 0.7, size = 8, family = 'serif'),
                    axis.text.y  = element_text(angle = 90, hjust = 0.5, family = "serif"),
                    axis.title   = element_text(family = "serif"),
                    plot.title   = element_text(family = "serif", face = "bold"),
                    legend.title = element_text(family = "serif", face = "bold"),
                    legend.text  = element_text(family = "serif")) +
              labs(title = "Levels of English Proficiency Within Different Income Brackets",
                    x = "Income Brackets",
                    y = "Percentage")


```

# Hypothesis 3: Income Bracket v.s. Religion
We predict that income brackets will vary based on religion and that individuals identifying as Hindu have the highest percentage of top earners on average. 

```{r, echo = FALSE, message=FALSE, warning=FALSE, error=FALSE}
IncomeBasedOnReligion <- AAQoL %>% 
  select(Income, Religion) %>% 
  na.omit() %>% 
  group_by(Income, Religion) %>% 
  summarize(n = n())

totalresp <- AAQoL %>% select(Income, Religion) %>% 
  na.omit() %>%
  group_by(Religion) %>%
  summarize(total_responses = n())

IncomeBasedOnReligionFinal <- merge(IncomeBasedOnReligion, totalresp, by = "Religion") %>% mutate(percentage = n / total_responses * 100)

#V1 with labels

ggplot(IncomeBasedOnReligionFinal, aes(x = Religion, y = percentage, fill = Income)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            position = position_stack(vjust = 0.5), # Adjust the position of labels
            size = 3, color = "white") +  # You can customize size and color of the labels
  labs(title = "Percentage of People in Each Religion Group by Income",
       x = "Income",
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

```


#Hypothesis 4: Employment Status v.s. Sense of Belonging

We predict that those who are employed full-time will have greater levels of belonging and those who are unemployed will have lower levels of belonging.

```{r, echo = FALSE, message=FALSE, warning=FALSE, error=FALSE}
BelongingAndFullTimeEmployment <- AAQoL %>% 
  select(Belonging, Full.Time.Employment, Part.Time.Employment, Student, Homemaker) %>% 
  na.omit() %>% mutate(Employment = if_else(Full.Time.Employment != "0", Full.Time.Employment, Part.Time.Employment)) %>% 
  mutate(Employment = if_else(Employment != "0", Employment, Student))  %>%
  mutate(Employment = if_else(Employment != "0", Employment, Homemaker)) %>%
  select(-Full.Time.Employment, -Part.Time.Employment, -Student, -Homemaker) %>% group_by(Belonging, Employment) %>%
  filter(Employment != 0)%>% 
  summarize(n = n())

totalbelong <- BelongingAndFullTimeEmployment %>%
  group_by(Belonging) %>%
  summarize(total_belonging = sum(n))



BelongingAndEmployment <- merge(BelongingAndFullTimeEmployment, totalbelong, by = "Belonging") %>% mutate(percentage = n / total_belonging * 100)

# Plot using ggplot
ggplot(BelongingAndEmployment, aes(x = Belonging, y = percentage, fill = Employment)) +
  geom_bar(stat = "identity") +
  labs(title = "Level of belonging based on employment",
       x = "Belonging",
       y = "Percentage") + geom_text(aes(label = sprintf("%.1f%%", percentage)),
                                     position = position_stack(vjust = 0.5), # Adjust the position of labels
                                     size = 3, color = "white") + 
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()
```

