---
title: "Asian American Quality of Life Project"
output:
  github_document:
---

# Links

- [Dataset Source](https://data.austintexas.gov/dataset/Final-Report-of-the-Asian-American-Quality-of-Life/hc5t-p62z)

# Explanation of Project and Information about Data

This dataset that we are examining involves the fastest growing minority group in the United States, Asian Americans. Within this dataset, the broad term includes individuals having origins in the Far East, Southeast Asia, or the Indian subcontinent. The Final Report of the Asian American Quality of Life (AAQoL) comes from data provided by the city of Austin, TX, and includes 2,609 rows of data in which each row is an individual survey. There are 231 columns in the dataset, each column providing a different piece of information about the respondents from a question on the survey they were provided. Our project is evaluating the quality of life of Asian Americans and factors that contribute to that. Our data set will show why the quality of life may change based upon different variables and why it is significant.

# Goals for Project

Some goals of our project include exploring what factors influence economic well-being in Asian American populations. We examine relationships between income and education, English proficiency and religion. Using visualizations and data modeling we would like to determine how these variables are related and if we can infer why see differences between groups.



Hypothesis 1:

Income Bracket v.s. Education vs. Ethnicity
We predict a variation in educational attainment amongst the different types of Asian ethnicities. We also expect to see a positive general relationship between education and income.
# Hypothesis 1 Visualization

```{R}
# Import the dataset into R environment.
library(ggplot2)
library(dplyr)
library(tidyverse)
library(readr)
library(ggthemes)
Final_Report_of_the_Asian_American_Quality_of_Life_AAQoL_ <- read_csv("Final_Report_of_the_Asian_American_Quality_of_Life__AAQoL_.csv")

# Load the dataset into a data frame named 'AAQoL'
AAQoL <- data.frame(Final_Report_of_the_Asian_American_Quality_of_Life_AAQoL_)

#Removing NAs
test3 <- AAQoL %>%
  filter(!is.na(Income) & !is.na(Education.Completed) & !is.na(Ethnicity))


#Graphing education on y-axis, income bracket on x-axis, colored by ethnicity.
new_colors <- c("orange", "red", "blue", "lightblue2", "maroon")

test3 %>% filter(Ethnicity != "Other", Student != "Student", Retired != "Retired", Unemployed != "Unemployed") %>%
  ggplot(aes(x = Income, y = Education.Completed, fill = Ethnicity)) +
  geom_bar(stat = "summary", fun = mean, position = "dodge", na.rm = TRUE) +
  geom_errorbar(stat="summary", fun.data=mean_se, width=.5, position = position_dodge(width = 0.9))+
  coord_flip() +
  coord_cartesian(ylim = c(10, 18)) +
  theme_economist_white() +
  theme(axis.text.x = element_text(angle = 50, vjust = 0.5, hjust = .4, size = 9)) +
  scale_fill_manual(values = new_colors) +
labs(x = "Household Income Bracket", y = "Mean Education Completed")
```

Hypothesis 2:

Income Bracket v.s. English Proficiency 
We predict that the level of English proficiency in the Asian American population will be positively related with their income bracket.
# Hypothesis 2 Visualization
```{R}

library(tidyverse)
library(extrafont)
loadfonts()


df_unfiltered <- read_csv("Final_Report_of_the_Asian_American_Quality_of_Life__AAQoL_.csv")
# Data Wrangling, 2609 observations to 1620 observations
df <- df_unfiltered %>%
  filter(`US Born` == 'No') %>% # Filter for only immigrants, 256 rows removed.
  filter(('Full Time Employment' == 'Employed full time')|('Part Time Employment' == 'Employed part time')|(Retired=='Retired')) %>%
  filter(Income != '') %>% # Filter for only response with income information, 167 rows removed
  filter('English Speaking' != '') %>% # Filter for only responses with proficiency information, 6 rows removed
  select(1,Income,'Proficiency'='English Speaking',21:22,Retired,Age) %>% # Select only relevant variables
  na.omit() # Omit all cells with N/A, 263 rows removed.
  # TO-DO
  # Ask TA About how to join the employment columns

# Redefining Proficiency as a ordinal variable
df$Proficiency <- factor(df$Proficiency, ordered = TRUE,
                         levels = c("Not at all", "Not well", "Well", "Very well"))

# Primary Visualization
df %>% ggplot(aes(x = Income, fill = Proficiency)) +
              geom_bar(position = 'fill') +
              scale_fill_brewer(palette = 15) +
              theme(axis.text.x  = element_text(angle = 15, hjust = 0.7, size = 8, family = 'serif'),
                    axis.text.y  = element_text(angle = 90, hjust = 0.5, family = "serif"),
                    axis.title   = element_text(family = "serif"),
                    plot.title   = element_text(family = "serif", face = "bold"),
                    legend.title = element_text(family = "serif", face = "bold"),
                    legend.text  = element_text(family = "serif")) +
              labs(title = "Levels of English Proficiency Within Different Income Brackets",
                    x = "Income Brackets",
                    y = "Percentage")

# Redefining Age and Income for the second visualization
df <- df %>%
  mutate(Income5 = recode(Income, "$10,000 - $19,999" = "$10,000 - $29,999",
                                  "$20,000 - $29,999" = "$10,000 - $29,999",
                                  "$30,000 - $39,999" = "$30,000 - $49,999",
                                  "$40,000 - $49,999" = "$30,000 - $49,999",
                                  "$50,000 - $59,999" = "$50,000 - $69,999",
                                  "$60,000 - $69,999" = "$50,000 - $69,999")) %>% # Recode Income into 5 categories
  mutate(AgeCat = case_when(Age >= 18 & Age <= 30 ~ "Age 18-30 (n=302)",
                            Age >= 31 & Age <= 40 ~ "Age 31-40 (n=344)",
                            Age >= 41 & Age <= 55 ~ "Age 41-55 (n=414)",
                            Age >= 55 ~ "Age 55+ (n=403)",
                            TRUE ~ NA_character_)) # Create a categorical variable for age

# Make sure each each gorup is evenly sized
sum(df$AgeCat == "Age 18-30 (n=302)")
sum(df$AgeCat == "Age 31-40 (n=344)")
sum(df$AgeCat == "Age 41-55 (n=414)")
sum(df$AgeCat == "Age 55+ (n=403)")


```

Hypothesis 3:

Income Bracket v.s. Religion
We predict that income brackets will vary based on religion and that individuals identifying as Hindu have the highest percentage of top earners on average. 
# Hypothesis 3 Visualization
```{r}
IncomeBasedOnReligion <- AAQoL %>% 
  select(Income, Religion) %>% 
  na.omit() %>% 
  group_by(Income, Religion) %>% 
  summarize(n = n())

totalresp <- AAQoL %>% select(Income, Religion) %>% 
  na.omit() %>%
  group_by(Religion) %>%
  summarize(total_responses = n())

IncomeBasedOnReligionFinal <- merge(IncomeBasedOnReligion, totalresp, by = "Religion") %>% mutate(percentage = n / total_responses * 100)

ggplot(IncomeBasedOnReligionFinal, aes(x = Religion, y = percentage, fill = Income)) +
  geom_bar(stat = "identity") +
  labs(title = "Percentage of People in Each Religion Group by Income",
       x = "Income",
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

#V1 with labels

ggplot(IncomeBasedOnReligionFinal, aes(x = Religion, y = percentage, fill = Income)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            position = position_stack(vjust = 0.5), # Adjust the position of labels
            size = 3, color = "white") +  # You can customize size and color of the labels
  labs(title = "Percentage of People in Each Religion Group by Income",
       x = "Income",
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()


#V2

IncomeBasedOnReligionFinal1 <- IncomeBasedOnReligionFinal

Income_level <- c("$70,000 and over", "$60,000 - $69,999", "$50,000 - $59,999", "$40,000 - $49,999", "$30,000 - $39,999", "$20,000 - $29,999", "$10,000 - $19,999", "$0 - $9,999")

IncomeBasedOnReligionFinal1 <- IncomeBasedOnReligionFinal1 %>% mutate(Income = factor(Income, levels = Income_level))

ggplot(IncomeBasedOnReligionFinal1, aes(x = Religion, y = percentage, fill = Income)) +
  geom_bar(stat = "identity") +
  labs(title = "Percentage of People in Each Religion Group by Income",
       x = "Income",
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()
```


Hypothesis 4:

Employment Status v.s. Sense of Belonging
We predict that those who are employed full-time will have greater levels of belonging and those who are unemployed will have lower levels of belonging.
#Hypothesis 4 Visualization
```{r}
BelongingAndFullTimeEmployment <- AAQoL %>% 
  select(Belonging, Full.Time.Employment, Part.Time.Employment, Student, Homemaker) %>% 
  na.omit() %>% mutate(Employment = if_else(Full.Time.Employment != "0", Full.Time.Employment, Part.Time.Employment)) %>% 
  mutate(Employment = if_else(Employment != "0", Employment, Student))  %>%
  mutate(Employment = if_else(Employment != "0", Employment, Homemaker)) %>%
  select(-Full.Time.Employment, -Part.Time.Employment, -Student, -Homemaker) %>% group_by(Belonging, Employment) %>%
  filter(Employment != 0)%>% 
  summarize(n = n())

totalbelong <- BelongingAndFullTimeEmployment %>%
  group_by(Belonging) %>%
  summarize(total_belonging = sum(n))



BelongingAndEmployment <- merge(BelongingAndFullTimeEmployment, totalbelong, by = "Belonging") %>% mutate(percentage = n / total_belonging * 100)

# Plot using ggplot
ggplot(BelongingAndEmployment, aes(x = Belonging, y = percentage, fill = Employment)) +
  geom_bar(stat = "identity") +
  labs(title = "Level of belonging based on employment",
       x = "Belonging",
       y = "Percentage") + geom_text(aes(label = sprintf("%.1f%%", percentage)),
                                     position = position_stack(vjust = 0.5), # Adjust the position of labels
                                     size = 3, color = "white") + 
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()
```

